---
title: "Reproducible Research - Project1"
author: "Lopamudra Satpathy"
date: "November 17, 2017"
output: html_document
---
#INTRODUCTION:
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.
Dataset: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
date: The date on which the measurement was taken in YYYY-MM-DD format
interval: Identifier for the 5-minute interval in which measurement was taken
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.
#Necessary packages to be installed
```{r}
library(lubridate)
library(dplyr)
library(reshape2)
library(ggplot2)

```

# Loading and pre processing the data
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
if(!file.exists("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip")) {
  temp <- tempfile()
  download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
  unzip(temp)
  unlink(temp)
}
activitydata <- read.csv("./activity.csv", sep = ",")
names(activitydata)
str(activitydata)
head(activitydata)
activitydata$date <- as.Date(activitydata$date)
```
# Mean of total No of steps taken per day (using reshape2 package)
Calculate the total number of steps taken per day
Histogram of total no of steps taken each day
Mean and Median of total no of steps taken per day
```{r}

actMeltDat <- melt(activitydata, id= "date",measure.vars = "steps", na.rm = TRUE)
actCastDat <-dcast(actMeltDat,date ~ variable,sum)
plot(actCastDat$date, actCastDat$steps, type="h", main="Histogram of Daily Steps", xlab="Date", 
     ylab="Steps per Day", col="blue", lwd= 8)
abline(h=mean(actCastDat$steps, na.rm=TRUE), col="red", lwd=2)
Mean <- mean(actCastDat$steps, na.rm = TRUE)
Median <- median(actCastDat$steps, na.rm = TRUE)

```
##What is the average daily activity pattern? (using package reshape2)
#Average daily pattern by interval
#Time series plot of the average number of steps taken (Average frequency of steps by interval )
```{r}
actMeltInt <- melt(activitydata, id="interval", measure.vars = "steps", na.rm = TRUE)
actCastInt <- dcast(actMeltInt, interval ~ variable, sum)
plot(actCastInt$interval, actCastInt$steps, type="l",main = "Frequeny of steps per interval",
     xlab = "Interval", ylab = "Steps", col= "purple", lwd = 3 )
abline(h = mean(actCastInt$steps, na.rm = TRUE),col = "blue", lwd = 2)
maxInterval <- actCastInt[which.max(actCastInt$steps),1]

```
##Strategy for imputing missing values
1.Calculate totall number of missing values
2.To fill the missing value,we choose to replace the value with the mean value.
3.We create the function na_fill(data,pervalue), in which data is the activitydata dataframe and pervalue is 
actCastInt dataframe.
4.Histogram of total no of steps taken each day
5.Number of rows with NA values
```{r}
sum(is.na(activitydata$steps))
naFill <- function(activitydata, pervalue) {
  naIndex <- which(is.na(activitydata$steps))
  naReplace <- unlist(lapply(naIndex, FUN=function(idx){
    interval = activitydata[idx,]$interval
    pervalue[pervalue$interval == interval,]$steps
  }))
  fillSteps <- activitydata$steps
  fillSteps[naIndex] <- naReplace
  fillSteps
}

activitydataFill <- data.frame(  
  steps = naFill(activitydata, actCastInt),  
  date = activitydata$date,  
  interval = activitydata$interval)
str(activitydataFill)
sum(is.na(activitydataFill$steps))
totalSteps <- aggregate(steps ~ date, activitydataFill,sum)
colnames(totalSteps) <- c("date", "steps")
hist(totalSteps$steps,totalSteps$date, main = "Histogram of total steps taken per day", 
     xlab = "Total steps per day", ylab = "Number of days", 
     breaks = 10, col = "steel blue")
abline(v = mean(totalSteps$steps), lty = 1, lwd = 2, col = "red")
abline(v = median(totalSteps$steps), lty = 2, lwd = 2, col = "black")
legend(x = "topright", c("Mean", "Median"), col = c("red", "black"), 
       lty = c(2, 1), lwd = c(2, 2))
sum(is.na(activitydata$steps))
sum(is.na(activitydata$steps))*100/nrow(activitydata)
```
#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
#compare the activities in weekday and weekend 
#For this I have used the dataset in which the missing data is filled in.
# Using dplyr,mutate to create new column dayIndicator
#Calculation of average steps interval in 5min 
#Using ggplot2 for the plotting of timeseries of 5min of weekday and weekend and compare the average steps
```{r}
activitydataFill <- mutate(activitydataFill, dayIndicator = 
                             ifelse(weekdays(activitydataFill$date) == "Saturday" | 
                                      weekdays(activitydataFill$date) == "Sunday", "weekend", "weekday"))
activitydataFill$dayIndicator <- as.factor(activitydataFill$dayIndicator)
head(activitydataFill) 
averageInterval <- activitydataFill %>%
  group_by(interval, dayIndicator) %>%
  summarise(steps = mean(steps))
g <- ggplot(averageInterval, aes(x=interval, y=steps, color = dayIndicator)) +
  geom_line() +
  facet_wrap(~dayIndicator, ncol = 1, nrow=2)
print(g)
```

